---
title: "Translation of Field's 'Meta_Basic_r.sps' script into R"
author: "Ian Hussey"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

Note that Field's script implements three meta analysis approaches: fixed effects, Hedges and colleagues method, and the Hunter-Schmidt method (without correlation deattenuations).

I translated only the H&S method below into R from Field's SPSS code.

```{r, include=FALSE}

knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE)

```

# Data and dependencies

```{r}

library(tidyverse)
library(janitor)

data_input <- read.csv("../../../data/data_vahey_et_al_2015.csv") %>%
  select(r_weighted_by_df, df_supplementary) %>%
  drop_na()

r <- data_input$r_weighted_by_df
n <- data_input$df_supplementary

```

# As implemented in Field's script

Only for Hunter-Schmidt RE model, not Hedges or fixed effects

```{r}

# get number of studies (k) from number of rows in r-vector
k = length(r)
# obtain adjusted values of r  (Overton, 1998, page 358)
ar = r - (r*(1-r^2))/(2*(n-3))     # unnumbered equation between equations 8 and 9 in Field & Gillett (2010)

# Hunter-Schmidt random-effects analysis
rav = sum(n*ar)/sum(n)
sr2 = sum(n*((ar-rav)^2))/sum(n)  # equation 2 in Field & Gillett (2010)
se2 = (1-rav^2)^2/((sum(n)/k)-1)  # equation 3 in Field & Gillett (2010)

# variance in population correlations 
vrho = sr2 - se2     # equation 4 in Field & Gillett (2010)

# NB Meta_Basic_r.sps does not contain this conditional to set SD to 0 if SE is <=0, but h_s_syntax.sps does. I've therefore added it here.
if(vrho > 0){
  sdrho = sqrt(vrho)
} else {
  sdrho = 0
}

cr_lower = rav - 1.96*sdrho   # equation 5 in Field & Gillett (2010)
cr_upper = rav + 1.96*sdrho

```

Sample Correlation Variance = `r round_half_up(sr2, 3)`

- Note that this differs from that reported by Vahey et al. (2015), i.e., .006.
  
Sampling Error Variance = `r round_half_up(se2, 3)`

Estimated Variance in Population Correlations (i.e., sr2 - se2) = `r vrho`

- Note that because this value is negative, confidence intervals cannot be calculated.

meta r = `r round_half_up(rav, 2)`, 95% CR [`r round_half_up(cr_lower, 2)`, `r round_half_up(cr_upper, 2)`]

# Removing Field's application of Overton's correction

I realised that Field & Gillett (2010), in defining the three different methods of meta-analysis that they discussed, state that *only* the Hedges' method involves Fisher's r-to-z transformations and backtransformations (equations 7 and 8), and the use of Overton's transformation prior to r-to-z to debias the z scores (unnumbered equation in text between equations 8 and 9). However, Field's implements of these methods (i.e., "Meta_Basic_r.sps" and the above translation of it into R) incorrectly also apply the Overton transformation to the correlations used in the Hunter & Schmidt style meta analysis (i.e., without using r-to-z transformations). Below, I therefore attempt to reproduce Vahey et al.'s (2015) results by modifying Field's translated code to remove these transformations.

```{r}

# Hunter-Schmidt random-effects analysis
rav_nooverton = sum(n*r)/sum(n)
sr2_nooverton = sum(n*((r-rav_nooverton)^2))/sum(n)  # equation 2 in Field & Gillett (2010)
se2_nooverton = (1-rav_nooverton^2)^2/((sum(n)/k)-1)  # equation 3 in Field & Gillett (2010)

# variance in population correlations 
vrho_nooverton = sr2_nooverton - se2_nooverton   # equation 4 in Field & Gillett (2010)

# NB Meta_Basic_r.sps does not contain this conditional to set SD to 0 if SE is <=0, but h_s_syntax.sps does. I've therefore added it here.
if(vrho_nooverton > 0){
  sdrho_nooverton = sqrt(vrho_nooverton)
} else {
  sdrho_nooverton = 0
}

cr_lower_nooverton = rav_nooverton - 1.96*sdrho_nooverton   # equation 5 in Field & Gillett (2010)
cr_upper_nooverton = rav_nooverton + 1.96*sdrho_nooverton

```

Sample Correlation Variance = `r round_half_up(sr2_nooverton, 3)`

- Note that this differs from that reported by Vahey et al. (2015), i.e., .006.
  
Sampling Error Variance = `r round_half_up(se2_nooverton, 3)`

Estimated Variance in Population Correlations (i.e., sr2 - se2) = `r vrho_nooverton`

- Note that because this value is negative, confidence intervals cannot be calculated.

meta r = `r round_half_up(rav_nooverton, 2)`, 95% CR [`r round_half_up(cr_lower_nooverton, 2)`, `r round_half_up(cr_upper_nooverton, 2)`]

