---
title: "Translation of Field's 'Meta_Basic_r.sps' script into R"
author: "Ian Hussey"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    code_folding: hide
    highlight: haddock
    theme: flatly
    toc: yes
    toc_float: yes
---

```{r, include=FALSE}

knitr::opts_chunk$set(message = FALSE,
                      warning = FALSE)

```

# Data and dependencies

```{r}

library(tidyverse)
library(janitor)

data_input <- read.csv("../../../data/data_vahey_et_al_2015.csv") %>%
  select(mean_r_forest_plot_numeric, n_forest_plot) %>%
  drop_na()

r <- data_input$mean_r_forest_plot_numeric
n <- data_input$n_forest_plot

```

# Translation of Field's script

Note that Field's script implements three meta analysis approaches: fixed effects, Hedges and colleagues method, and the Hunter-Schmidt method (without correlation deattenuations).

I translated only the H&S method below into R from Field's SPSS code.

## As implemented in Field's script

Only for Hunter-Schmidt RE model, not Hedges or fixed effects

```{r}

# get number of studies (k) from number of rows in r-vector
k = length(r)
# obtain adjusted values of r  (Overton, 1998, page 358)
ar = r - (r*(1-r^2))/(2*(n-3))

# # create zr, Fisher's z-transformation of r
# zr = 0.5*log((1+r)/(1-r))
# # zr has variance = 1/(n-3)  (Field, 2001, page 164)
# v = 1/(n-3)
# # create fixed-effects weight vector
# w = 1/v
# # sum weights
# sw = sum(w)
# # calculate c  (Hedges & Vevea, 1998, equation 10)
# c = sw - t(w)*w/sw
# # calculate Q  (H & V equation 7)
# q = t(w)*zr^2 - (t(w)*zr)^2/sw
# # variance component tau is set equal to zero if Q < k-1,  and to  (q-k+1)/c  otherwise
# tau = max(c(0, (q-k+1)/c))
# # variance of effect size under random-effects model
# vt = v + tau
# # random-effects weight vector (H & V equation 13)
# wt = 1/vt
# # sum weights
# swt = sum(wt)
# # calculate Q for R-E
# qe = t(wt)*zr^2 - (t(wt)*zr)^2/swt

# # RE:
# mean = t(wt)*zr/swt
# 
# # calculate standard errors and confidence intervals for means
# sem = sqrt(1/swt)
# 
# ci = c(mean - 1.96*sem, mean + 1.96*sem)
# # convert means and confidence intervals (based on zr) back to r
# rmean = (exp(2*mean)-1)/(exp(2*mean)+1)
# #rci = (exp(2*ci)-1)/(exp(2*ci)+1) # possibly incorrectly converted
# zscore = abs(mean/sem)
# pz = 2*(1 - pnorm(zscore))

# Hunter-Schmidt random-effects analysis
rav = sum(n*ar)/sum(n)
sr2 = sum(n*((ar-rav)^2))/sum(n)
se2 = (1-rav^2)^2/((sum(n)/k)-1)

# variance in population correlations Field & Gillett (2010) equation 4
vrho = sr2-se2

# NB Field did not add a check for positive sdrho but I add one here
# sdrho = sqrt(vrho)
# cil = rav - 1.96*sdrho
# ciu = rav + 1.96*sdrho

if(vrho > 0){
  sdrho = sqrt(vrho)
  cil = rav - 1.96*sdrho
  ciu = rav + 1.96*sdrho
} else {
  cil = NA
  ciu = NA
}

# chi = sum((n-1)*(ar-rav)^2)/(1-rav^2)^2
# chisig = 1 - pchisq(chi, k-1)

# print('Chi2')
# chi
# print('p')
# chisig
# print('df')
# (k-1)

```

Sample Correlation Variance = `r round_half_up(sr2, 3)`

- Note that this differs from that reported by Vahey et al. (2015), i.e., .006.
  
Sampling Error Variance = `r round_half_up(se2, 3)`

Estimated Variance in Population Correlations (i.e., sr2 - se2) = `r vrho`

- Note that because this value is negative, confidence intervals cannot be calculated.

meta r = `r round_half_up(rav, 2)`, 95% CI [`r round_half_up(cil, 2)`, `r round_half_up(ciu, 2)`]

## Removing Field's application of Overton's correction

I realised that Field & Gillett (2010), in defining the three different methods of meta-analysis that they discussed, state that *only* the Hedges' method involves Fisher's r-to-z transformations and backtransformations (equations 7 and 8), and the use of Overton's transformation prior to r-to-z to debias the z scores (unnumbered equation in text between equations 8 and 9). However, Field's implements of these methods (i.e., "Meta_Basic_r.sps" and the above translation of it into R) incorrectly also apply the Overton transformation to the correlations used in the Hunter & Schmidt style meta analysis (i.e., without using r-to-z transformations). Below, I therefore attempt to reproduce Vahey et al.'s (2015) results by modifying Field's translated code to remove these transformations.

```{r}

# Hunter-Schmidt random-effects analysis
rav_nooverton = sum(n*r)/sum(n)
sr2_nooverton = sum(n*((r-rav_nooverton)^2))/sum(n)
se2_nooverton = (1-rav_nooverton^2)^2/((sum(n)/k)-1)

# variance in population correlations Field & Gillett (2010) equation 4
vrho_nooverton = sr2_nooverton-se2_nooverton

if(vrho_nooverton > 0){
  sdrho_nooverton = sqrt(vrho_nooverton)
  cil_nooverton = rav - 1.96*sdrho_nooverton
  ciu_nooverton = rav + 1.96*sdrho_nooverton
} else {
  cil_nooverton = NA
  ciu_nooverton = NA
}

```

Sample Correlation Variance = `r round_half_up(sr2_nooverton, 3)`

- Note that this differs from that reported by Vahey et al. (2015), i.e., .006.
  
Sampling Error Variance = `r round_half_up(se2_nooverton, 3)`

Estimated Variance in Population Correlations (i.e., sr2 - se2) = `r vrho_nooverton`

- Note that because this value is negative, confidence intervals cannot be calculated.

meta r = `r round_half_up(rav_nooverton, 2)`, 95% CI [`r round_half_up(cil_nooverton, 2)`, `r round_half_up(ciu_nooverton, 2)`]

# Additions

## Confidence interval

Confidence intervals are more commonly calculated (i.e., in non psychometric meta analyses) as simply estimate ± 1.96 * se. I try that here on the basis that perhaps Vahey et al defaulted to it when Field's script (above) could not calculate CIs by the H&S method. NB these use the non-overton transformed values.

### Using Overton correction

```{r}

cil_se = rav - 1.96*sqrt(se2)
ciu_se = rav + 1.96*sqrt(se2)

```

95% CR [`r round_half_up(cil_se, 2)`, `r round_half_up(ciu_se, 2)`]

### Without Overton correction

```{r}

cil_nooverton_se = rav_nooverton - 1.96*sqrt(se2_nooverton)
ciu_nooverton_se = rav_nooverton + 1.96*sqrt(se2_nooverton)

```

95% CR [`r round_half_up(cil_nooverton_se, 2)`, `r round_half_up(ciu_nooverton_se, 2)`]

## Credibility interval

Vahey et al. (2015) reported 95% CR [.23, .67].

Whereas the "h_s syntax.sps" file calculates a credibility interval, the "Meta_Basic_r.sps" file does not. 

However, such an interval can be calculated using the same formula as in the "h_s syntax.sps" file.

### Using Field's "h_s syntax.sps" formula

#### With Overton correction 

```{r}

# CR using s_r^2 calculated above 
crl = rav - 1.96*sqrt(sr2)
cru = rav + 1.96*sqrt(sr2)

```

95% CR [`r round_half_up(crl, 2)`, `r round_half_up(cru, 2)`]

#### Without Overton correction 

```{r}

# CR using s_r^2 calculated above
crl_nooverton = rav_nooverton - 1.96*sqrt(sr2_nooverton)
cru_nooverton = rav_nooverton + 1.96*sqrt(sr2_nooverton)

```

95% CR [`r round_half_up(crl_nooverton, 2)`, `r round_half_up(cru_nooverton, 2)`]

### Attempt to reproduce Vahey et al.'s (2015) 95% CR from their reported $s_r^2$

As a third attempt at verification, I too one step back and tried to reproduce their 95% CR from their reported $s_r^2$ (0.006) and $\overline{r}$ (.45).

```{r}

# s_r^2 (0.006) and r_bar (.45) as reported in Vahey et al. 2015
crl_verified = .45 - 1.96*sqrt(0.006)
cru_verified = .45 + 1.96*sqrt(0.006)

```

95% CR [`r round_half_up(crl_verified, 2)`, `r round_half_up(cru_verified, 2)`]

Following Field & Gillet's (2010) definition of a CR (equation 5), which is consistent with Field's implementation of the CR (see "h_s syntax.sps"), which Vahey reported to me in an email that he employed, Vahey et al.'s (2015) reported CR cannot be reproduced from their $s_r^2$ and $\overline{r}$.

